---
title: Homework 4. Question 3.4.
output: pdf_document
---


```{r message=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(knitr)
```


```{r fig.width = 4, fig.height = 3}
set.seed(20201112)
generatex <- function(n, mu, beta,N){
    # each experiment is column
    epsilon <- matrix(rnorm(n*N), ncol = N)
    
    e <- matrix(rep_len(0.0,(n+1)*N), ncol = N)
    X <- matrix(rep_len(0.0,n*N), ncol = N)
    i <- 2L
    for(i in seq(2,n+1)){
        e[i,] <- beta*e[i-1,] + epsilon[i-1,]
        X[i-1,] <- mu + e[i,]
    }
    epsilon[1,]
    X
}
mle_mu <- function(X) {
    X[1,]
}
mle_beta <- function(X) {
    (apply(X,2,mean)-X[1,])/(apply(X[1:(nrow(X)-1L),],2,mean)-X[1,])
}
mu <- 0.5
beta <- 0.7
X <- generatex(7,mu,beta,5)
X_df <- cbind(data.frame(i = seq(1, nrow(X))), as.data.frame(X))
ggplot(tidyr::gather(X_df,"id","X",-i), aes(i,X,color=id)) + geom_line()
```


```{r fig.width = 4, fig.height = 3}
set.seed(20201112)
mu <- 0.5
beta <- 0.7
X <- generatex(50,mu,beta,5000)
mu_est <- mle_mu(X)
beta_est <- mle_beta(X)

hist(mu_est)
hist(beta_est, breaks=20000L, xlim=c(0,2))
```

From plots ML estimate mu look promising to be normal, estimate for beta is spikier and with long tails. In addition to that estimator for beta looks significantly biased.

```{r}
shapiro.test(mu_est)
shapiro.test(beta_est)
```

Shapiro-Wilk normality test supports findings.

```{r}
set.seed(20201112)
mu <- 0.5
beta <- 0.7
cat(paste("mu=",mu,"beta=",beta,"\n"))
for(n in c(5,7,10,12,25,50)){
    X <- generatex(n,mu,beta,5000)
    mu_est <- mle_mu(X)
    beta_est <- mle_beta(X)
    mu_est_normtest <- shapiro.test(mu_est)
    beta_est_normtest <- shapiro.test(beta_est)
    cat(paste(
        "n=",n,
        "mu est norm p-value=", mu_est_normtest$p.value,
        "beta est norm p-value=",beta_est_normtest$p.value,"\n"))
}
```

ML estimate mu is normal, which is not surprising given that X[1]=mu+epsilon[1], i.e. it is normal.

ML estimate for beta failed Shapiro-Wilk normality test with p-value < 0.05 for all n, thus it is not normal.
